---
title: "Untitled"
author: "Tony Mei"
date: "11/21/2020"
output: html_document
---


```{r echo=FALSE}
# load required packages
library(ggplot2)
library(dplyr)
library(corrplot)
library(MASS)
library(caret)
library(RCurl)
library(pROC)
library(RCurl)
library(haven)
```

```{r import}
# Loading the data
git_dir <- 'https://raw.githubusercontent.com/odonnell31/DATA621-HW4/main/data'
#class_data = read.csv(paste(git_dir, "/classification-output-data.csv", sep=""))
train_df = read.csv(paste(git_dir, "/insurance_training_data.csv", sep=""))
test_df = read.csv(paste(git_dir, "/insurance-evaluation-data.csv", sep = ""))
nrow(train_df)
```

## Data Exploration & Preparation

See a summary of each column in the train_df set
```{r train_dfing_data_summary}
# view a summary of all columns
summary(train_df)
```


Look at the data type of each variable
```{r}
# data type of predictors
str(train_df)
```

Look at the relationship between TARGET_FLAG and some of the numerical variables.
```{r}
par(mfrow=c(1,2))
# plot response variable "target" against predictor variable "age" and "car_age"
boxplot(AGE ~ TARGET_FLAG, train_df, 
        main="Target vs Age",
        xlab="Target",
        ylab="Age") 
boxplot(CAR_AGE ~ TARGET_FLAG, train_df, 
        main="Target vs Car Age",
        xlab="Target",
        ylab="Car Age")
```

Look at the distribution of some numerical variables.
```{r}
h <- hist(train_df$AGE)
text(h$mids,h$counts,labels=h$counts)
```

```{r}
h <- hist(train_df$MVR_PTS)
text(h$mids,h$counts,labels=h$counts)
```


```{r}
numeric = function(input) {
  out = sub("\\$", "", input)
  out = as.numeric(sub(",", "", out))
  return(out)
}


train_df = as.tbl(train_df) %>% 
  mutate_at(c("INCOME","HOME_VAL","BLUEBOOK","OLDCLAIM"),
            numeric)

#check data
summary(train_df)

```

```{r}
train_df$AGE[is.na(train_df$AGE)] <- mean(train_df$AGE, na.rm=TRUE)
train_df$YOJ[is.na(train_df$YOJ)] <- mean(train_df$YOJ, na.rm=TRUE)
train_df$HOME_VAL[is.na(train_df$HOME_VAL)] <- mean(train_df$HOME_VAL, na.rm=TRUE)
train_df$CAR_AGE[is.na(train_df$CAR_AGE)] <- mean(train_df$CAR_AGE, na.rm=TRUE)

train_df$INCOME[is.na(train_df$INCOME)] <- mean(train_df$INCOME, na.rm=TRUE)

#get complete cases
train_df <- train_df[complete.cases(train_df),]

train_df2<-train_df

train_df2<-train_df2[-c(1)]


```


```{r}
train_amt_df <- subset(train_df, TARGET_AMT > 0)
summary(train_amt_df$TARGET_FLAG)
```

Binary Logistic Regression Models
```{r}
#model 1, preliminary exploration with one predictor

model1 <- glm(formula = TARGET_FLAG ~ AGE, family = binomial(), data = train_df2)
summary(model1)
```

```{r}
#model 2, includes all predictor variables
model2 <- glm(formula = TARGET_FLAG ~ .- TARGET_AMT, family = binomial(), data = train_df2)
summary(model2)
```


```{r}
step_BLR = stepAIC(model2)
summary(model2)
```

```{r}
# model 3, selcted significant variables from model 2
model3 = glm(formula = TARGET_FLAG ~ KIDSDRIV + INCOME + PARENT1 + HOME_VAL + MSTATUS + 
    EDUCATION + JOB + TRAVTIME + CAR_USE + BLUEBOOK + TIF + CAR_TYPE + 
    OLDCLAIM + CLM_FREQ + REVOKED + MVR_PTS + URBANICITY, family = binomial(), data = train_df2)
summary(model3)
```
### 2) Multiple Linear Regression

```{r}
#linear model1, with all predictors
lm1 = lm(TARGET_AMT ~ ., data = train_df2)
summary(lm1)
```

```{r}
step <- stepAIC(lm1, direction="both")
```


```{r}
#linear model2, selected
lm2 = lm(TARGET_AMT ~ TARGET_FLAG + PARENT1 + SEX + BLUEBOOK + REVOKED + 
    MVR_PTS + CAR_AGE, data = train_df2)
summary(lm2)
```
```{r}
library(car)
car::vif(model2)
```


Using model2 because of the lower AIC out of the three models
```{r}
test_df$AGE[is.na(test_df$AGE)] <- mean(test_df$AGE, na.rm=TRUE)
test_df$YOJ[is.na(test_df$YOJ)] <- mean(test_df$YOJ, na.rm=TRUE)
test_df$HOME_VAL[is.na(test_df$HOME_VAL)] <- mean(test_df$HOME_VAL, na.rm=TRUE)
test_df$CAR_AGE[is.na(test_df$CAR_AGE)] <- mean(test_df$CAR_AGE, na.rm=TRUE)

test_df$INCOME[is.na(test_df$INCOME)] <- mean(test_df$INCOME, na.rm=TRUE)

#get complete cases
test_df <- test_df[complete.cases(test_df),]

test_df2<-test_df

test_df2<-test_df2[-c(1)]

numeric = function(input) {
  out = sub("\\$", "", input)
  out = as.numeric(sub(",", "", out))
  return(out)
}


test_df2 = as.tbl(test_df2) %>% 
  mutate_at(c("INCOME","HOME_VAL","BLUEBOOK","OLDCLAIM"),
            numeric)
```


```{r}
# step_BLR prediction on test
test_df<-test_df[-c(1)]
test_preds_BLR = round(predict(model2, newdata=test_df, type='response'))
test_df$TARGET_FLAG = test_preds_BLR
test_preds_MLR = predict(lm2, newdata=test_df)
test_df$TARGET_AMT = test_preds_MLR
# write out evaluation data with predictions
write.csv(test_df2, 'eval_with_preds.csv')

```