---
title: "HW1"
author: "Team 1"
date: "September 13, 2020"
output: 
  prettydoc::html_pretty:
    theme: cayman
    toc: true
---

# Library

```{r}
# load required packages
library(ggplot2)
library(dplyr)
#library(tidyr)
library(corrplot)
library(MASS)
library(caret)
library(RCurl)
```

```{r import}
# Loading the data

git_dir <- 'https://raw.githubusercontent.com/Sizzlo/Data621/main'
train_df = read.csv(paste(git_dir, "/crime-training-data_modified.csv", sep=""))
test_df = read.csv(paste(git_dir, "/crime-evaluation-data_modified.csv", sep = ""))
head(train_df)
```

# Data Exploration & Preparation

See a summary of each column in the train_df set
```{r train_dfing_data_summary}
# view a summary of all columns
summary(train_df)
```

```{r}
# Correlations 
cor_train = cor(train_df,  use = "na.or.complete")
corrplot(cor_train)
```

```{r}
pairs(~ target + zn + indus
      + chas + nox + rm + age + dis + rad + tax + ptratio + lstat + medv, data = train_df)
```
```{r}
# look at the structure of the variables
str(train_df)
str(test_df)
```

The summary() function for the training and testing data sets indicates that there are no missing values in the data. 
The response variable "target" is binary with 1 indicates crime rate is above median cirme rate and 0 indicates crime rate is not above median crime rate. 

Let's observe how the target variable is effected by other factors:
1. The plot of "target" against "age" shows target equalling one (above median crime rate) increases as the proportion of owner-occupied units built prior to 1940 increaases; the boxplot further shows that a larger mean of proportions of owner-occupied units built prior to 1940 is assoicated with higher crime rate.
2. Plots of crime rate against pupil-teacher ratio indicate higher crime rate "1" is associated with higher pupil-teacher ratio.

```{r}
par(mfrow=c(2,2))
# plot response variable "target" against predictor variable "age" 
plot(train_df$age,train_df$target)
boxplot(age ~ target, train_df )

# plot response variable "target" against predictor variable "ptratio"
plot(train_df$ptratio,train_df$target)
boxplot(ptratio ~ target, train_df)
```


```{r}
# preliminary exploration glm models
glm(formula = target ~ age, family = binomial(), data = train_df)
glm(formula = target ~ ptratio , family = binomial(), data = train_df)
```